### 2024-04-02 paper/paper.qmd
initial commit

---
title: "From Touches to Goals: Modeling the Efficacy of Possession in Football"
subtitle: "Evaluating Poisson vs. Negative Binomial Models in Predicting Home Team Success"
author: 
  - Siqi Fei
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "This study delves into the intricate relationship between a football team's ball possession, as indicated by the number of times players touch the ball, and the team's scoring success. Through rigorous statistical analysis, we compared Poisson and negative binomial regression models to assess which better predicts the final score based on possession data. Our findings revealed a nuanced connection where more ball possession does not always equate to higher scores, challenging common football assumptions. The implications of this research extend beyond sports analytics, providing insights into the complex dynamics of performance metrics and their interpretation, thereby enriching the strategic understanding of team sports."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Football is the world's most popular sport, with a vast audience and professionals across the globe. Germany, with its reputation for football excellence, hosts the Bundesliga every year. Bundesliga is a league that draws global attention due to its high level of competition and skill. The data from this league is highly regarded for the precision and dependability, making it ideal for in-depth analysis. In this study, we examine data from the Bundesliga season 2023-2024 to better understand the outcomes of matches, particularly how ball possession relates to scoring goals.

This study uses data from Kaggle for the year 2024 [@bundesliga_dataset_2024], applies Poisson and negative binomial regression models to closely examine the link between how often home team players have the ball and the number of goals they score. The aim is to see how keeping the ball influences the team's ability to score.

Echoing Maher's article [@Maher1982], which instead of rejecting the Poisson model for football scores, further explored it by considering how well teams attack and defend, our study revisits these models with newest Bundesliga data. We're comparing the Poisson and the Negative Binomial models to figure out which one better reflects the number of goals in current football games. Through our analysis with the latest data, we aim to uncover which model truly captures the scoring trends in football today

Our paper outlines the importance of analyzing match data, how we did our study, what we found, and why it's important. We show how our findings help us better understand football and its scores. Our research offers new insights for coaches, players, and anyone interested in the science behind the sport.

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-04-19 paper/paper.qmd
renamed model,shiny app,scripts

@@
 ---
-title: "From Touches to Goals: Modeling the Efficacy of Possession in Football"
-subtitle: "Evaluating Poisson vs. Negative Binomial Models in Predicting Home Team Success"
+title: "Economics and Crime Patterns: The Interplay Between CPI Changes and Vehicle Theft Frequencies"
+subtitle: "Insights from Predictive Analytics and Visual Data Exploration"
 author: 
   - Siqi Fei
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: <https://github.com/FXXFERMI/Inflation-and-Crime-Patterns.git>. Link to the Shiny App:  <https://siqi-fei.shinyapps.io/VehicleTheft_Inflation/>"
 date: today
 date-format: long
-abstract: "This study delves into the intricate relationship between a football team's ball possession, as indicated by the number of times players touch the ball, and the team's scoring success.
-Through rigorous statistical analysis, we compared Poisson and negative binomial regression models to assess which better predicts the final score based on possession data.
-Our findings revealed a nuanced connection where more ball possession does not always equate to higher scores, challenging common football assumptions.
-The implications of this research extend beyond sports analytics, providing insights into the complex dynamics of performance metrics and their interpretation, thereby enriching the strategic understanding of team sports."
+abstract: "In this study, we researched the monthly incidence of vehicle thefts and inflation from 2014 to 2024 by using data sourced from the Toronto Police Service and Statistics Canada.
+We analyzed the data distribution and characteristics of the two datasets and combined both to conduct a correlation analysis by using Bayesian multiple linear regression model and a Shiny application.
+We found a significant positive correlation between the overall CPI and vehicle theft rates.
+This finding suggests that higher inflation may be associated with an increased number of vehicle theft.
+This study highlights the importance of economic indicators in predicting criminal behaviour which in terms provide valuable insights for policymakers and economic analysts."
 format: pdf
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 
+#### Workspace setup ####
 library(tidyverse)
-library(palmerpenguins)
+library(here)
+library(modelsummary)
+library(tidyr)
+library(reshape2)
+library(corrplot)
+library(lubridate)
+library(dplyr)
+library(knitr)
+library(rstanarm)
+library(bayesplot)
+
+#### Load Data ####
+# originally imported via 01-data_cleaning.R file
+
+cleaned_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
+merge_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
+raw_data_cpi <- read.csv(here::here("data/analysis_data/analysis_data_CPI.csv"))
+vehicle_theft_summary <- read.csv(here::here("data/analysis_data/Draft_analysis_data_Vehicle.csv"))
+raw_data_auto <- read.csv(here::here("data/raw_data/Auto_Theft_Open_Data.csv"))
+
+model_matrix <- model.matrix(Total_Vehicle_Thefts ~ Recreation_education_and_reading + All_items + 
+                                    Goods + Household_operations_furnishings_equipment + 
+                                    Health_and_personal_care +
+                                    Transportation + 
+                                    Shelter + Services + Food + Alcoho_tobacco_cannabis +
+                                    Energy + Gasoline + Clothing_footwear, data = cleaned_data)
+
 ```
@@
 
-Football is the world's most popular sport, with a vast audience and professionals across the globe.
-Germany, with its reputation for football excellence, hosts the Bundesliga every year.
-Bundesliga is a league that draws global attention due to its high level of competition and skill.
-The data from this league is highly regarded for the precision and dependability, making it ideal for in-depth analysis.
-In this study, we examine data from the Bundesliga season 2023-2024 to better understand the outcomes of matches, particularly how ball possession relates to scoring goals.
+In recent years, rapid global economic development has accelerated currency devaluation.
+During the pandemic, constrained economic output had led to significant job losses and a shortage in global supply chains.
+These factors have contributed to a sharp rise in inflation rates across Canada.
+Data from @statcanCPI indicates that these changes have made it increasingly difficult for people to afford their daily expenses.
+Meanwhile, regional crime rates have also significantly increased.
+According to the Canada Newswire @equite2023, auto theft trends in Ontario rose dramatically by 48.2%, with a vehicle now stolen every five minutes between 2021 and 2023.
 
-This study uses data from Kaggle for the year 2024 [@bundesliga_dataset_2024], applies Poisson and negative binomial regression models to closely examine the link between how often home team players have the ball and the number of goals they score.
-The aim is to see how keeping the ball influences the team's ability to score.
+@McIntoshLi2012 highlight a critical gap in understanding the economic efficiency of crime prevention programs within the Canadian context, emphasizing the need for evidence-based approaches in crime prevention.
 
-Echoing Maher's article [@Maher1982], which instead of rejecting the Poisson model for football scores, further explored it by considering how well teams attack and defend, our study revisits these models with newest Bundesliga data.
-We're comparing the Poisson and the Negative Binomial models to figure out which one better reflects the number of goals in current football games.
-Through our analysis with the latest data, we aim to uncover which model truly captures the scoring trends in football today
+This study aims to fill this gap by examining the impact of Consumer Price Index (CPI) changes on vehicle theft rates over a ten-year period, from January 2014 to February 2024.
+Employing Bayesian multiple linear regression model, we analyzed the trends and variances in CPI across different sectors and correlated these economic indicators with the frequency of auto thefts in Toronto.
+The datasets were sourced from the Toronto Police Service @torontoPoliceAutoTheft2023 and Statistics Canada @statscanCPI2023.
+Our findings indicate that while an increase in the overall cost of living correlates with higher vehicle theft rates, the rise in Shelter CPI appears to inversely affect theft frequencies.
+Interestingly, other CPI categories such as food, transportation, health, and personal care did not show significant statistical influence on theft rates.
 
-Our paper outlines the importance of analyzing match data, how we did our study, what we found, and why it's important.
-We show how our findings help us better understand football and its scores.
-Our research offers new insights for coaches, players, and anyone interested in the science behind the sport.
+The estimand of this study is the effect of changes in the CPI on vehicle theft rates, which is quantitatively explored through our models.
+This research provides valuable contributions in understanding how economic conditions affect crime rates which provides important insights for policymakers and economic analysts.
+By integrating data visualization and interactive tools via a Shiny app, this paper enhances user engagement and facilitates deeper exploration of the economic variables influencing crime rates.
+
+This paper is structured as follows: @sec-data introduces the datasets utilized for the analysis, offering insights into the variables of interest and their visual representations.
+@sec-model details the Bayesian multiple linear regression model applied to explore the relationship between different CPI indicators' change and vehicle theft rates.
+@sec-results presents the results and interpretation of the models, and @sec-discussion discusses the implications of these findings, the limitations of the current study, and directions for future research.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Data Sources
+
+The databases used for this study primarily focus on the Consumer Price Index (CPI) in Canada and vehicle theft incidents in the Toronto area.
+We have utilized open data from @statscanCPI2023 and the @torontoPoliceAutoTheft2023.
+Statistics Canada released the latest update in April 2024 and the data set uses 2002 as the base year to record monthly changes in Canada’s CPI from January 2014 to February 2024 covering 15 different categories.
+The Toronto Police Service’s Auto Theft Open Data spans from January 2014 to March 2024, detailing each vehicle theft incident with its occurrence time and location.
+
+While other datasets were considered, the CPI data from Statistics Canada was chosen for its authoritative and wide coverage, and the Toronto Police Service's data was selected due to its reliability and detailed record of incidents over a decade.
+These sources provide the most relevant and reliable data for analyzing the relationship between economic conditions and crime rates.
+
+The analysis of this paper makes use of the R programming language [@citeR] for statistical computations and visualizing data.
+The tidyverse package [@citetidyverse] is installed to gain access to other important R packages, including the dplyr package [@citedplyr] used to manipulate and clean data, the readr package [@citereadr] to read and import data, the here package [@citehere] to create a path to specific saved files.
+The ggplot2 package [@citeggplot2], reshape2 package [@citereshape2], corrplot package [@citecorrplot] and lubridate package [@citelubridate] are used to create the data visualizations.
+And the modelsummary package [@citemodelsummary] to create summary tables.
+
+## Canada Consumer Price Index (CPI) Dataset
+
+### Sampling Strategy and Survey Methodology
+
+The Consumer Price Index (CPI) data set from Statistics Canada @statscanCPI2023, a sample represented in [Appendix -@sec-dataset-details], utilizes a multi-stage sampling strategy.
+It targets both urban and rural private households across Canada, while excluding non-representative groups such as inmates or individuals in collective households.
+This probability-based approach ensures a diverse demographic representation.
+
+Monthly price data collection relies on detailed specifications for a wide array of goods and services.
+This method allows for accurate tracking of price changes over time, using data from retail outlets and service providers.
+This broad item coverage accurately reflects Canadian consumer spending behaviours.
+
+### Variable of Interest
+
+For this study, we selected 13 major CPI categories:
+
+-   `All Items`: an aggregate measure of overall inflation.
+-   `Recreation,Education and Reading`: costs related to leisure activities, covering educational expenses and reading materials.
+-   `Goods`: including consumer products like electronics and clothing.
+-   `Household Operations, Furnishings and Equipment`: expenses for maintaining a household.
+-   `Health and Personal Care`: encompassing medical and health-related expenses.
+-   `Transportation`: expenses related to personal and public transport.
+-   `Shelter`: costs associated with housing.
+-   `Services`: various personal and professional services.
+-   `Food`: expenses on groceries and dining out.
+-   `Alcohol, Tobacco and Cannabis`: spending on alcoholic drinks, tobacco, and cannabis.
+-   `Energy`: household energy expenses.
+-   `Gasoline`: specifically fuel costs.
+-   `Clothing and Footwear`: spending on personal attire.
+
+All CPI data points are numeric and indexed to the base year of 2002, set at 100.
+
+## Toronto Vehicle Theft Dataset
+
+### Sampling Strategy and Survey Methodology
+
+The Toronto Police Service's dataset on auto theft @torontoPoliceAutoTheft2023 occurrences, a sample represented in [Appendix -@sec-dataset-details], is structured at the offense and vehicle level.
+Each reported occurrence can have multiple records, reflecting different Major Crime Indicators (MCIs) used to categorize each event.
+Data include the report date (REPORT_DATE) and the date the offense occurred (OCC_DATE), both standardized to UTC timezone to ensure uniformity.
+For privacy protection, exact locations of crimes are offset to the nearest road intersection, and the data excludes occurrences deemed unfounded—where an investigation determined that the reported offense did not occur or was not attempted.
+
+The auto theft dataset is updated quarterly, reflecting the dynamic nature of crime reporting.
+Additionally, all historical data ranges are provided to ensure coverage of trends over time.
+This dataset is intended to enhance public safety awareness by making detailed crime data accessible to the community while adhering to stringent privacy and data integrity standards.
+
+### Variable of Interest
+
+We have compiled this database to extract monthly vehicle theft frequencies, which serve as the primary data for this research.
+Featuring columns for `Year_Month` and `Total_Thefts` reflects a unique month-year combination and the corresponding count of vehicle thefts.
+The sample of compiled dataset represented below in @tbl-CompiledAutoData.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: tbl-CompiledAutoData
+#| tbl-cap: Sample of Compiled Vehicle Thetfs Dataset
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+selected_data <- vehicle_theft_summary[, c(2, 3)]
+kable(head(selected_data, 5), caption = "Sample of Compiled Vehicle Thetfs Dataset")
 ```
 
-Talk more about it.
+## Data Measurement
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+To ensure a coherent analysis, we synchronized the two datasets by aligning them according to common temporal identifiers, specifically month and year.
+The @tbl-analysisData is an example of final analysis dataset after merge CPI dataset and auto thefts dataset.
+This method allowed us to accurately correlate changes in the Consumer Price Index (CPI) with instances of vehicle thefts across the same time periods.
+The dataset includes 14 variables and 122 monthly observations from January 2014 through February 2024.
+
+Before analysis, we conducted a review of the datasets to ensure their integrity. This process included:
+
+-   Verification of Data Completeness: Although both datasets were initially found to be free of missing values and outliers, we performed additional checks to confirm the completeness and accuracy of all 122 observations.
+In cases where missing values were detected, we replaced them with the data from the preceding month.
+
+-   Consistency and Alignment Checks: We cross-validated the monthly CPI values against the reported vehicle thefts to ensure that each observation was correctly matched and no temporal misalignments were present.
+
+By adhering to these data verification protocols, our study ensures a high level of data quality, enabling a focused and reliable examination of how various economic conditions as depicted by CPI data correlate with trends in vehicle thefts across the Toronto area.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: tbl-analysisData
+#| tbl-cap: Sample of Analysis Dataset
+#| echo: false
+
+selected_data <- cleaned_data[, c(1, 7, 3, 6, ncol(cleaned_data))]
+kable(head(selected_data), caption = "Sample of Analysis Dataset")
+```
+
+## Data Distribution
+
+### Canada Consumer Price Index (CPI)
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+#### Plot Set Up ####
+# Pivot the data to long format for plotting
+data_long <- pivot_longer(cleaned_data, 
+                          cols = c(Recreation_education_and_reading, Goods, Household_operations_furnishings_equipment,
+                                   Health_and_personal_care, All_items, Transportation, Shelter, Services, Food,
+                                   Alcoho_tobacco_cannabis, Energy, Gasoline, Clothing_footwear),
+                          names_to = "Category", 
+                          values_to = "Index")
+
+merge_data_filtered <- cleaned_data[, !names(cleaned_data) %in% c("All_items", "Energy")]
+```
+
+::: {#fig-CPIDistribution}
+![](images/box_all_cpi.png){width="73%"}
+
+Box Plots for Distribution - all CPI categories
+:::
+
+The @fig-CPIDistribution shows the range of Consumer Price Index (CPI) values for 13 categories, illustrating the variability in price changes from 2014 to 2024.
+Some categories, like 'Alcohol, Tobacco, and Cannabis,' display wider ranges, indicating more pronounced fluctuations in prices over the years.
+Others, such as 'Clothing and Footwear,' have narrower boxes, suggesting steadier prices.
+
+The median of each category, marked by a line in the box, provides a quick reference for the middle value of the data.
+Notably, several categories feature outliers, which point to occasional extremes in pricing that could be explored further.
+
+This box plot allows us to observe the distribution patterns of CPI values, serving as a basis for analyzing economic trends within these categories.
+
+### Toronto Vehicle Theft
+
+```{r}
+#| label: fig-Distribution_Thefts
+#| fig-cap: Distribution of Vehicle Thefts from 2014 to 2024
 #| echo: false
@@
 #| message: false
+# Distribution Plot
+# Set up the plotting area
+par(mar = c(5, 5, 4, 2) + 0.1)
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+# Create the histogram with a density curve
+hist(merge_data$Total_Vehicle_Thefts, 
+     breaks = 20, 
+     freq = FALSE, 
+     main = "Distribution of Vehicle Thefts", 
+     xlab = "Total Vehicle Thefts",
+     ylab = "Density",
+     col = "lightblue",
+     xlim = c(0, max(merge_data$Total_Vehicle_Thefts) * 1.1))
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# Add a density curve
+lines(density(merge_data$Total_Vehicle_Thefts), col = "darkred", lwd = 2)
+
+# Add a legend
+legend("topright", 
+       legend = c("Density Curve"), 
+       col = c("darkred"), 
+       lwd = c(2), 
+       bty = "n")
+
+# Add a rug plot to show individual data points
+rug(merge_data$Total_Vehicle_Thefts, col = "darkblue", lwd = 0.5)
+
 ```
 
-Talk way more about it.
+The @fig-Distribution_Thefts with an overlaid density curve illustrates the distribution of total vehicle thefts.
+Most theft incidents cluster in the lower range of the scale, indicating a higher frequency of months with fewer thefts.
+The right-skewed distribution suggests that while most months have a moderate number of thefts, there are occasional months with significantly higher occurrences.
+The tail of the histogram, where it extends towards larger values, reflects these rarer, higher-theft months.
+The density curve highlights the overall shape of the distribution, emphasizing the concentration of data points around the mode.
 
-# Model
+```{r}
+#| label: tbl-Distribution_Thefts
+#| tbl-cap: Summary statistics table of Total Vehicle Thefts from 2014 to 2024
+#| echo: false
+#| warning: false
+#| message: false
+#| 
+summary_table <- merge_data %>%
+  summarise(
+    Mean = mean(Total_Vehicle_Thefts, na.rm = TRUE),
+    Median = median(Total_Vehicle_Thefts, na.rm = TRUE),
+    SD = sd(Total_Vehicle_Thefts, na.rm = TRUE),
+    Min = min(Total_Vehicle_Thefts, na.rm = TRUE),
+    Max = max(Total_Vehicle_Thefts, na.rm = TRUE),
+    Q1 = quantile(Total_Vehicle_Thefts, 0.25, na.rm = TRUE),
+    Q3 = quantile(Total_Vehicle_Thefts, 0.75, na.rm = TRUE),
+    IQR = IQR(Total_Vehicle_Thefts, na.rm = TRUE),
+    Count = n()
+  )
 
-The goal of our modelling strategy is twofold. Firstly,...
+kable(summary_table, caption = "Descriptive Statistics of Total Vehicle Thefts")
+```
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The @tbl-Distribution_Thefts offers a direct insight into the distribution of total vehicle thefts over the study period.
+The mean thefts per month stand at 495.21, with a median lower at 421, indicating a skew in the data with some months experiencing higher theft counts.
+The standard deviation is substantial at 253.36, suggesting a wide variability in monthly theft incidents.
+
+The data range spans from a minimum of 201 thefts in a month to a maximum of 1164, revealing extreme fluctuations that could reflect varying external factors influencing theft rates.
+The first quartile (Q1) and the third quartile (Q3) are at 309 and 607.25 thefts respectively, with an interquartile range (IQR) of 298.25, which highlights the middle 50% of the data is less variable compared to the full range.
+
+# Model {#sec-model}
+
+The goal of our modeling strategy is twofold.
+Firstly, we aim to evaluate the impact of various Consumer Price Index (CPI) categories on the rates of vehicle thefts in Toronto.
+We explore which economic factors are significant predictors of vehicle theft occurrences.
+Secondly, we seek to provide estimates of the effects of these factors, quantifying how significant changes in CPI influence vehicle theft statistics.
+
+Here we briefly describe the Bayesian multiple linear regression model used to investigate the relationships between economic conditions and vehicle thefts.
+Background details are included in [Appendix -@sec-model-details].
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+We define $y_i$ as the monthly count of vehicle thefts. The model includes thirteen Consumer Price Index (CPI) categories, each represented by a coefficient $\beta_1$ to $\beta_{13}$.
 
+```{=tex}
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_1 \cdot \text{All\_items}_i + \beta_2 \cdot \text{Recreation\_education\_and\_reading}_i + \\
+&\quad + \beta_3 \cdot \text{Goods}_i + \beta_4 \cdot \text{Household\_operations\_furnishings\_equipment}_i + \\
+&\quad + \beta_5 \cdot \text{Health\_and\_personal\_care}_i + \beta_6 \cdot \text{Transportation}_i + \\
+&\quad + \beta_7 \cdot \text{Shelter}_i + \beta_8 \cdot \text{Services}_i + \beta_9 \cdot \text{Food}_i + \\
+&\quad + \beta_{10} \cdot \text{Alcohol\_tobacco\_cannabis}_i + \beta_{11} \cdot \text{Energy}_i + \beta_{12} \cdot \text{Gasoline}_i + \\
+&\quad + \beta_{13} \cdot \text{Clothing\_footwear}_i \\
+\alpha &\sim \text{Normal}(0, 2.5) \\
+\beta_1, \beta_2, \ldots, \beta_{13} &\sim \text{Normal}(0, 2.5) \\
+\sigma &\sim \text{Exponential}(1)
 \end{align}
+```
+We run the Bayesian model in R [@citeR] using the `rstanarm` package of @rstanarm.
+We apply Normal priors with a mean of zero and standard deviation of 2.5 are applied to each of the coefficients, providing a balance between allowing the data to inform the model and maintaining prior constraints to avoid overly flexible interpretations.
+The error variance $\sigma$ is assumed to follow an exponential distribution, emphasizing the assumption of non-negative error variability.
+This approach captures the nuanced effects of economic indicators on vehicle theft rates, modeling each category's potential impact separately.
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+## Model justification
 
-### Model justification
+We expect a notable relationship between the various CPI categories and the rates of vehicle thefts.
+In particular, economic factors such as $\beta_1$ (All_items), $\beta_6$ (Transportation), and $\beta_{12}$ (Gasoline) may significantly influence theft incidents due to their impact on the cost of living and mobility.
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+Given the economic context, categories like $\beta_9$ (Food) and $\beta_{10}$ (Alcohol_tobacco_cannabis) might also influence crime rates, as changes in these indices often reflect shifts in the broader economy that could affect criminal behaviour.
+The inclusion of $\beta_4$ (Household_operations_furnishings_equipment) and $\beta_{13}$ (Clothing_footwear) is intended to capture the effects of economic conditions on consumer vulnerability to theft.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+# Results {#sec-results}
 
-# Results
+::: {#fig-TimeSeries}
+![](images/Time.png)
 
-Our results are summarized in @tbl-modelresults.
+Time Series Plot for all CPI categories and Vehicle Thefts from 2014 to 2024
+:::
+
+The @fig-TimeSeries illustrates the evolution of Consumer Price Index (CPI) categories alongside vehicle thefts from January 2014 through February 2024.
+Each line represents a different CPI category or the vehicle thefts, providing a visual representation of trends over time.
+
+-   The category of Total Vehicle Thefts (in pink) exhibits a notable increase starting around early 2021, reaching a peak in mid-2022 before a downward trend.
+This rise and subsequent fall may correlate with external factors affecting vehicle theft incidents, which warrants further investigation.
+
+-   The 'All_items' CPI category shows relative stability until it starts to rise around the same time as vehicle thefts, suggesting a possible relationship between overall price levels and theft incidents.
+
+-   Other CPI categories such as 'Health_and_personal_care', 'Clothing_footwear', and 'Household_operations_furnishings_equipment' demonstrate a stable trend with slight fluctuations, indicating less variability in these categories over the observed period.
+
+-   The 'Energy' and 'Gasoline' categories exhibit some volatility, reflecting market changes that could influence consumer behaviour and potentially crime rates.
+
+-   Interestingly, while most CPI categories remain within a tighter band, the 'Total Vehicle Thefts' category shows more pronounced movement, especially during the later years of the series.
+This could be indicative of underlying economic or social shifts impacting theft rates more significantly than changes in consumer prices.
+
+The model results are summarized in @tbl-modelresults.
 
@@
 
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+bayesian_linear_model <-
+  readRDS(file = here::here("models/bayesian_linear_model.rds"))
 ```
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| tbl-cap: "Explanatory models of number of Vehicle Thefts based on CPI"
 #| warning: false
@@
   list(
-    "First model" = first_model
+    "Bayesian Linear Model" = bayesian_linear_model
   ),
@@
 
-# Discussion
+Our analysis presents a robust relationship between CPI categories and vehicle thefts.
+The model, adjusting for multiple variables, offers a nuanced understanding of how different economic indicators influence theft rates.
 
-## First discussion point {#sec-first-point}
+-   The intercept at -3458.80 establishes a baseline for the model; without any CPI influence, the model predicts a significantly negative number of thefts, which isn't possible in reality and suggests the model's reliance on the CPI predictors.
+-   The 'All_items' category shows a positive coefficient (52.25), indicating that a unit increase in this category is associated with an increase of approximately 52 vehicle thefts, signifying the strong effect of overall cost of living on theft incidents.
+-   Interestingly, 'Recreation_education_and_reading' shows a negative coefficient (-22.20), suggesting that higher expenditure in this category may be associated with a decrease in theft rates, possibly reflecting discretionary spending's correlation with reduced crime rates.
+-   'Goods' has a small positive effect (11.43), while 'Household_operations_furnishings_equipment' has a negative coefficient (-5.59), which might indicate that as people invest more in their homes, the rate of thefts slightly declines.
+-   'Services' stands out with a significant positive coefficient (50.08), highlighting the potential impact of service expenditures on theft rates, a factor that urban planners and policymakers may need to consider.
+-   'Shelter' has a negative coefficient (-32.67), which could imply that more spending on housing correlates with lower theft rates, a factor of interest in urban development.
+-   The coefficients for 'Food' (-0.33) and 'Health_and_personal_care' (0.77) are relatively small, indicating a less pronounced direct impact on vehicle theft rates.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The model's R-squared value of 0.938 suggests a very high proportion of variance in theft rates is explained by these CPI variables, signaling a strong fit.
+The RMSE of 292.30 underscores the model's precision, with predictions on average about 292 units from the observed values.
 
-## Second discussion point
+# Discussion {#sec-discussion}
 
-## Third discussion point
+## Overview
 
-## Weaknesses and next steps
+This paper examined how changes in the Consumer Price Index (CPI) relate to vehicle theft rates.
+Our analysis considered multiple CPI categories to identify which aspects of economic change have an effect on crime.
 
-Weaknesses and next steps should also be included.
+## Insights on Pandemic
+
+The pandemic brought significant changes to the Consumer Price Index (CPI), showcasing the economic shifts caused by global lockdowns and altered consumer behaviors.
+Notably, the 'All_items' CPI increased due to heightened demand and supply chain interruptions, escalating the cost of daily goods and services.
+Price fluctuations in the 'Energy' and 'Gasoline' sectors were marked by sharp volatility, reflecting decreased global demand and the impacts of oil price conflicts.
+Conversely, expenditures on 'Recreation_education_and_reading' remained stable, likely due to a shift in spending towards home-based activities during public restrictions.
+
+Simultaneously, the pandemic's economic stress led to an increase in vehicle thefts, mirroring the surge in CPI.
+This rise particularly impacted categories critical for daily living, suggesting that economic pressures may drive individuals towards criminal activities as a coping mechanism.
+Changes in public movement and lockdown measures reshaped traditional crime patterns, indicating shifts in the timing and locations of criminal activities.
+
+This period of profound economic uncertainty and enforced social restrictions clearly demonstrates the interconnectedness of economic factors and crime rates over time.
+It underscores the necessity for vigilant, responsive social and economic strategies during crises to effectively mitigate the escalation of crime, reflecting the enduring impact of such global events on societal norms and safety.
+
+## Insights on Economic Conditions and Crime Dynamics
+
+This study also explores the connection between Consumer Price Index (CPI) categories and vehicle theft rates in Canada over a decade.
+It finds that increases in the 'All_items' CPI, reflecting the overall cost of living, align with more vehicle thefts.
+However, spending in areas like recreation and education correlates with fewer thefts, suggesting discretionary spending may deter crime.
+
+The relationship between vehicle thefts and essential goods like 'Health_and_personal_care' and 'Clothing_footwear' appears less direct.
+Meanwhile, 'Energy' and 'Gasoline' prices show unique patterns, indicating market fluctuations might influence theft rates differently.
+
+By examining CPI categories separately, the study highlights how economic conditions affect crime.
+It offers insights that could guide policies for economic management and crime prevention, showing how various economic pressures impact crime differently and underscoring the role of economic analysis in improving societal conditions.
+
+## Limitations
+
+This study aimed to explore the impact of economic changes on crime rates. However, it faces several limitations:
+
+1.
+**Scope of Economic Factors:** The research focused on CPI and vehicle thefts but did not include other potentially influential economic factors such as unemployment rates or income disparity.
+Expanding the range of economic indicators could provide a broader understanding of what drives crime rates.
+
+2.
+**Geographic Limitation:** The findings are based on data from Toronto and may not apply to other regions with different socio-economic conditions.
+Future studies could analyze data from multiple regions to draw more general conclusions.
+
+3.
+**Population Bias:** As a major urban center, Toronto's large population might influence the study results, potentially making them less applicable to smaller or rural areas.
+Including a more varied set of locations could help balance this bias.
+
+4.
+**Impact of the Pandemic:** The study acknowledges the pandemic's effects but does not analyze them in depth.
+Further research could more closely examine how specific aspects of the pandemic influenced economic and crime trends.
+
+5.
+**Data Reliability:** The study relies on reported crime data and CPI, which could be subject to underreporting or inconsistencies in data collection.
+More rigorous data verification could strengthen the findings.
+
+6.
+**Causal Pathways:** The study does not investigate how economic changes directly lead to variations in crime rates.
+Future research might include qualitative approaches to better understand the causal relationships involved.
+
+## Future Directions
+
+Given the limitations identified in the study, here are what we can do in the future study:
+
+1.
+**Broader Economic Indicators**: Future studies should incorporate a wider range of economic indicators, including unemployment rates, income levels, and economic policy changes, to more fully understand their impact on crime rates.
+
+2.
+**Expanded Geographic Scope**: To address the geographic limitation, subsequent research could include data from multiple cities and rural areas across different countries.
+This would help determine if the findings from Toronto are applicable elsewhere or if local conditions significantly influence crime dynamics.
+
+3.
+**Diverse Demographic Analysis**: Investigating how economic changes affect different demographic groups could uncover nuanced patterns of how economic conditions influence crime rates among various populations.
+
+4.
+**In-depth Pandemic Analysis**: A focused study on the pandemic period could clarify how temporary economic disruptions and social restrictions specifically affected crime rates, providing insights into managing crime during crises.
+
+5.
+**Enhanced Data Collection and Verification**: To improve data reliability, future research should use more rigorous methods for data collection and verification.
+This could involve cross-referencing crime data with other sources or using advanced statistical techniques to handle potential data inaccuracies.
+
+6.
+**Causal Research**: Implementing mixed-methods approaches, including qualitative research, could help elucidate the causal pathways between economic conditions and crime, offering a clearer picture of the underlying mechanisms.
+
+7.
+**Policy Impact Studies**: Researching the effects of specific economic and social policies on crime rates could provide valuable feedback for policymakers, helping to design interventions that effectively address the economic drivers of crime.
+
+By pursuing these directions, future research can build on the current study’s findings and address its shortcomings, leading to a deeper understanding of the complex relationship between economics and crime.
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
-# Additional data details
+# Additional dataset details {#sec-dataset-details}
+
+The @tbl-CPIData shows a sample dataset of CPI from @statscanCPI2023. The @tbl-AutoData shows a sample dataset of Vehicle Thetfs Dataset from @torontoPoliceAutoTheft2023.
+
+```{r}
+#| label: tbl-CPIData
+#| tbl-cap: Sample of CPI Dataset
+#| echo: false
+
+selected_data <- raw_data_cpi[, c(1, 7, 3, 6)]
+kable(head(selected_data, 15), caption = "Sample of CPI Dataset")
+```
+
+```{r}
+#| label: tbl-AutoData
+#| tbl-cap: Sample of Vehicle Thetfs Dataset
+#| echo: false
+
+selected_data <- raw_data_auto[, c(5, 6, 12, 13)]
+kable(head(selected_data), caption = "Sample of Vehicle Thetfs Dataset")
+```
 
@@
 
+## Shiny App - Data Visualization
+
+We developed a Shiny application to facilitate interactive exploration of the dataset derived from our study.
+The application, built with the R packages `shiny` [@citeShiny], `ggplot2` [@citeggplot2], and `DT` [@citeDT], allows users to dynamically view the relationships between various Consumer Price Index (CPI) categories and vehicle theft occurrences.
+Upon selection of a CPI category from the sidebar, the main panel updates to display a scatter plot illustrating the correlation with vehicle thefts, complemented by a regression line to indicate trends.
+The tool also features a downloadable data option, enabling users to export the currently viewed data subset for further analysis.
+This interactive approach promotes a deeper engagement with the data and supports additional hypothesis generation based on visual insights.
+
+The @fig-LPAll-items shows a sample scenario in the shiny app. The shiny app can be found through link: <https://siqi-fei.shinyapps.io/VehicleTheft_Inflation/>
+
+```{r}
+#| label: fig-LPAll-items
+#| fig-cap: Relationship between All-items CPI and Vehicle Thefts
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 10
+#| fig-height: 5
+ggplot(cleaned_data, aes(x = All_items, y = Total_Vehicle_Thefts)) +
+  geom_point() + # This adds the scatter plot points
+  geom_smooth(method = "lm", se = FALSE) + # This adds a regression line without a confidence interval
+  labs(title = "Relationship between All-items CPI and Vehicle Thefts",
+       x = "All-items CPI",
+       y = "Total Vehicle Thefts") +
+  theme_minimal() # This applies a minimal theme to the plot
+```
+
+```{r}
+#| label: fig-LPServices
+#| fig-cap: Relationship between Services CPI and Vehicle Thefts
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: false
+
+ggplot(cleaned_data, aes(x = Services, y = Total_Vehicle_Thefts)) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE) + 
+  labs(title = "Relationship between Services CPI and Vehicle Thefts",
+       x = "All-items CPI",
+       y = "Total Vehicle Thefts") +
+  theme_minimal() 
+```
+
+```{r}
+#| label: fig-LPGoods
+#| fig-cap: Relationship between Goods CPI and Vehicle Thefts
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: false
+
+ggplot(cleaned_data, aes(x = Goods, y = Total_Vehicle_Thefts)) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE) + 
+  labs(title = "Relationship between Goods CPI and Vehicle Thefts",
+       x = "All-items CPI",
+       y = "Total Vehicle Thefts") +
+  theme_minimal() 
+```
+
+```{r}
+#| label: fig-LPRecreation_education_and_reading
+#| fig-cap: Relationship between Recreation, Education and Reading CPI and Vehicle Thefts
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: false
+
+ggplot(merge_data_filtered, aes(x = Recreation_education_and_reading)) +
+  geom_point(aes(y = Total_Vehicle_Thefts)) +
+  geom_smooth(aes(y = Total_Vehicle_Thefts), method = "lm", se = FALSE) +
+  facet_wrap(~., scales = "free", ncol = 2) +
+  labs(title = "Relationship between CPI Categories (excluding All-items and Energy) and Vehicle Thefts",
+       x = "CPI Category",
+       y = "Total Vehicle Thefts")
+
+```
+
 ## Posterior predictive check
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check.
+This shows how closely the simulated data aligns with our observed data, and highlighting the model's predictive accuracy.
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
+This shows the influence of the observed data on our initial assumptions and the resulting adjustments to our parameter estimates.
 
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
+pp_check(bayesian_linear_model, type = "hist", nsamples = 100)
 
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
+bayesplot::mcmc_areas(
+  as.array(bayesian_linear_model),
+  pars = c("All_items", "Recreation_education_and_reading", "Goods", "Household_operations_furnishings_equipment", "Health_and_personal_care", "Transportation", "Shelter","Services", "Food", "Alcoho_tobacco_cannabis", "Energy","Gasoline", "Clothing_footwear"), # specify your parameters
+  prob = 0.95
+)
 
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-04-19 Vehicle_Theft_Shiny_App/app.R

@@ -1,0 +1,74 @@
+#### Work space Setup ####
+library(shiny)
+library(ggplot2)
+library(DT)
+
+# Load dataset
+transposed_data <- read.csv("analysis_data_copy.csv")
+
+# Define the user interface
+ui <- fluidPage(
+  titlePanel("Interactive Data Viewer"),
+  sidebarLayout(
+    sidebarPanel(
+      selectInput("xAxisColumn", "Select CPI Category for X-axis:",
+                  choices = colnames(transposed_data)[-which(colnames(transposed_data) == "Total_Vehicle_Thefts")],
+                  selected = "All_items"),
+      hr(),
+      downloadButton("downloadData", "Download Data")
+    ),
+    mainPanel(
+      DTOutput("dataTable"),
+      hr(),
+      plotOutput("dataPlot")
+    )
+  )
+)
+
+# Define server logic
+server <- function(input, output) {
+  # Reactive expression to handle the data
+  filteredData <- reactive({
+    # Ensure selected column exists in transposed_data
+    if (input$xAxisColumn %in% colnames(transposed_data)) {
+      transposed_data[, c(input$xAxisColumn, "Total_Vehicle_Thefts"), drop = FALSE]
+    } else {
+      NULL  # Return NULL if the column does not exist
+    }
+  })
+  
+  # Render the DataTable
+  output$dataTable <- renderDT({
+    if (is.null(filteredData())) {
+      return(dataTableOutput())  # Return an empty table if data is NULL
+    }
+    datatable(filteredData())
+  })
+  
+  # Downloadable csv of selected dataset
+  output$downloadData <- downloadHandler(
+    filename = function() {
+      paste("data-", Sys.Date(), ".csv", sep="")
+    },
+    content = function(file) {
+      write.csv(filteredData(), file, row.names = FALSE)
+    }
+  )
+  
+  # Dynamic scatter plot with regression line
+  output$dataPlot <- renderPlot({
+    if (is.null(filteredData())) {
+      return()  # Do nothing if data is NULL
+    }
+    req(filteredData())  # ensure that data is not NULL
+    ggplot(filteredData(), aes_string(x = input$xAxisColumn, y = "Total_Vehicle_Thefts")) +
+      geom_point() +
+      geom_smooth(method = "lm", se = FALSE) +
+      labs(title = paste("Relationship between", input$xAxisColumn, "CPI Inflation Rate and Vehicle Thefts Times"),
[…]

### 2024-04-19 scripts/00-simulate_data.R

@@ -21,39 +21,60 @@ library(tidyverse)
 # Set seed for reproducibility
 set.seed(42)
 
-# Create a sequence of months and years from January 2014 to March 2024
+# Create a sequence of months and years from January 2014 to Feb. 2024
 months <- rep(1:12, times = 10)  # 12 months for each year (2014 to 2023)
-months <- c(months, 1:3)  # Add first three months of 2024
-print(months)
+months <- c(months, 1:2)  # Add first two months of 2024
 years <- rep(2014:2023, each = 12)  # 10 years (2014 to 2023)
-print(years)
+years <- c(years, rep(2024, 2)) # two year rows for 2024
+
+#print(years)
+#print(months)
+
+# Simulate data for Auto Theft Dataset
+
+# Initialize an empty vector to store increasing values
+increasing_values <- numeric(length(months))
+
+# Set the initial value
+initial_value <- 250
+
+# Generate increasing values with random increments
+for (i in 1:length(months)) {
+  random_increment <- sample(5:15, 1)  # Random increment between 5 and 15
+  increasing_values[i] <- initial_value + random_increment * i
+}
+
 
 # Simulate data for Auto Theft Dataset
 auto_theft <- data.frame(
   month = months,
   year = years,
-  Total_number = sample(100:500, size = length(months), replace = TRUE)
+  Total_Vehicle_Thefts = increasing_values
 )
 
 # Simulate data for CPI Dataset
+initial_cpi <- 100  # Initial CPI value in 2002
+
 cpi <- data.frame(
   month = months,
   year = years,
-  Food = runif(length(months), 100, 200),
-  Shelter = runif(length(months), 120, 220),
-  Clothing_and_footwear = runif(length(months), 80, 180),
-  Gasoline = runif(length(months), 90, 190),
-  Health_and_personal_care = runif(length(months), 100, 200),
-  Goods = runif(length(months), 110, 210),
-  Household_operations_furnishings_and_equipment = runif(length(months), 130, 230),
-  Alcoholic_beverages_tobacco_products_and_recreational_cannabis = runif(length(months), 140, 240),
-  Transportation = runif(length(months), 70, 170),
-  Recreation_education_and_reading = runif(length(months), 90, 190),
[…]

### 2024-04-19 scripts/01-data_cleaning.R

@@ -175,6 +175,7 @@ merge_data <- merge_data %>%
 write_csv(merge_data, "data/analysis_data/analysis_data.csv")
 write_csv(raw_data_cpi, "data/analysis_data/analysis_data_CPI.csv")
 write_csv(vehicle_theft_wide, "data/analysis_data/analysis_data_Vehicle.csv")
+write_csv(vehicle_theft_summary, "data/analysis_data/draft_analysis_data_Vehicle.csv")
 
 
 #### Additional Clean ####
@@ -197,8 +198,19 @@ transposed_data <- transposed_data[-1, ]
 transposed_data <- tibble::rownames_to_column(transposed_data, var = "Month_Year")
 transposed_data[,-1] <- lapply(transposed_data[,-1], as.numeric)
 
+
+# Remove columns under headers All_items_excluding_food_and_energy and All_items_excluding_energy
+transposed_data <- subset(transposed_data, select = -c(All_items_excluding_food_and_energy, All_items_excluding_energy))
+
+# Get the index of the column with header Total_Vehicle_Thefts
+col_index <- which(names(transposed_data) == "Total_Vehicle_Thefts")
+
+# Move the column to the third position
+transposed_data <- transposed_data[, c(setdiff(seq_along(transposed_data), col_index), col_index)]
+
 # Check the structure
 #str(transposed_data)
 
 #### Save data ####
 write_csv(transposed_data, "data/analysis_data/analysis_data.csv")
+

### 2024-04-19 scripts/02-test_data.R

@@ -23,7 +23,7 @@ library(testthat)
 library(tidyverse)
 
 #### Load Data ####
-merge_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
+transposed_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
 raw_data_cpi <- read.csv(here::here("data/analysis_data/analysis_data_CPI.csv"))
 vehicle_theft_wide <- read.csv(here::here("data/analysis_data/analysis_data_Vehicle.csv"))
 
@@ -60,17 +60,17 @@ test_that("Vehicle theft data has correct number of columns", {
 
 # Test that the merged dataset has the correct dimensions
 test_that("Merged data has correct dimensions", {
-  expect_equal(nrow(merge_data), 122)
-  expect_equal(ncol(merge_data), 17)
+  expect_equal(nrow(transposed_data), 122)
+  expect_equal(ncol(transposed_data), 15)
 })
 
 # Test that the merged dataset has no NA values
 test_that("Merged data has no NA values", {
-  expect_true(all(complete.cases(merge_data)))
+  expect_true(all(complete.cases(transposed_data)))
 })
 
 # Test that all columns in CPI data are numeric after conversion
 test_that("CPI data contains only numeric columns", {
-  expect_true(all(sapply(merge_data[-1], is.numeric)))
+  expect_true(all(sapply(transposed_data[-1], is.numeric)))
 })
 

### 2024-04-19 scripts/03-data_measurement.R

@@ -1,0 +1,133 @@
+#### Preamble ####
+# Purpose: Prepares and cleans the dataset for analysis, including handling missing values, 
+#          normalizing data, and creating derived variables for further analysis.
+# Author: Siqi Fei
+# Date: 05 April 2024
+# Contact: fermi.fei@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: This script requires 'tidyverse' for data manipulation and 'lubridate' for 
+#                 date handling. Ensure that the raw data is in the specified directory.
+# Additional Notes: This script generates a cleaned dataset saved in a standard CSV format for use in modeling.
+
+#### Set Up ###
+#install.packages("tidyr")
+#install.packages("corrplot")
+#install.packages("lubridate")
+#install.packages("reshape2")
+library(tidyr)
+library(reshape2)
+library(corrplot)
+library(lubridate)
+
+#### Load Data ####
+merge_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
+raw_data_cpi <- read.csv(here::here("data/analysis_data/analysis_data_CPI.csv"))
+vehicle_theft_wide <- read.csv(here::here("data/analysis_data/analysis_data_Vehicle.csv"))
+
+#### CPI Data Measurements ####
+
+# Pivot the data to long format for plotting
+data_long <- pivot_longer(merge_data, 
+                          cols = c(Recreation_education_and_reading, Goods, Household_operations_furnishings_equipment,
+                                   Health_and_personal_care, All_items, Transportation, Shelter, Services, Food,
+                                   Alcoho_tobacco_cannabis, Energy, Gasoline, Clothing_footwear),
+                          names_to = "Category", 
+                          values_to = "Index")
+
+# Scatter Plot with Regression Line
+ggplot(merge_data, aes(x = All_items, y = Total_Vehicle_Thefts)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(title = "Relationship between All-items CPI and Vehicle Thefts",
+       x = "All-items CPI",
+       y = "Total Vehicle Thefts")
+
+ggplot(merge_data, aes(x = Energy, y = Total_Vehicle_Thefts)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(title = "Relationship between Energy CPI and Vehicle Thefts",
[…]

### 2024-04-19 scripts/03-model.R

@@ -1,54 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Siqi Fei
-# Date: 05 April 2024
-# Contact: fermi.fei@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-#install.packages("rstanarm")
-#install.packages("tidyverse")
-#install.packages("rstan")
-library(tidyverse)
-library(rstanarm)
-library(rstan)
-
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
-
-### Model data ####
-# Linear regression model including all CPI categories
-linear_model <- lm(Total_Vehicle_Thefts ~ Recreation_education_and_reading + 
-                     Goods + Household_operations_furnishings_equipment + 
-                     All_items_excluding_food_and_energy + Health_and_personal_care +
-                     All_items_excluding_energy + All_items + Transportation + 
-                     Shelter + Services + Food + Alcoho_tobacco_cannabis +
-                     Energy + Gasoline + Clothing_footwear,
-                   data = analysis_data)
-
-
-# Summarize the model to view coefficients and statistics
-summary(linear_model)
-
-# Diagnostic plots for the regression model
-par(mfrow = c(2, 2))
-plot(linear_model)
-
-# If not already installed
-# install.packages("stargazer")
-library(stargazer)
-
-# Create a nice HTML or text table of your model's results
-stargazer(linear_model, type = "text")
-
-
-
-
-#### Save model ####
-#saveRDS(
-# first_model,
-#file = "models/first_model.rds"
-#)

### 2024-04-19 scripts/04-data_measurements.R

@@ -1,42 +1,0 @@
-# should be removed to qmd later
-#### Set Up ###
-#install.packages("corrplot")
-#install.packages("reshape2")
-library(reshape2)
-library(corrplot)
-#### Load Data ####
-merge_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
-raw_data_cpi <- read.csv(here::here("data/analysis_data/analysis_data_CPI.csv"))
-vehicle_theft_wide <- read.csv(here::here("data/analysis_data/analysis_data_Vehicle.csv"))
-
-#### CPI Data Measurements ####
-
-# Scatter Plot with Regression Line
-ggplot(merge_data, aes(x = All_items, y = Total_Vehicle_Thefts)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  labs(title = "Relationship between All-items CPI and Vehicle Thefts",
-       x = "All-items CPI",
-       y = "Total Vehicle Thefts")
-
-# Box Plots for Distribution
-# Using melt from the data.table package to reshape data for plotting
-long_format <- melt(merge_data, id.vars = "Month_Year", variable.name = "CPI_Category", value.name = "CPI_Value")
-
-ggplot(long_format, aes(x = CPI_Category, y = CPI_Value)) +
-  geom_boxplot() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs(title = "Distribution of CPI by Category",
-       x = "CPI Category",
-       y = "CPI")
-
-
-#### Vehicle Theft Data Measurements ####
-
-# Distribution Plot
-hist(merge_data$Total_Vehicle_Thefts, breaks = 30, main = "Histogram of Vehicle Thefts", xlab = "Total Vehicle Thefts")
-
-# Trend Analysis
-trend_model <- lm(Total_Vehicle_Thefts ~ Time, data = merge_data)
-summary(trend_model)
-

### 2024-04-19 scripts/04-model.R

@@ -1,0 +1,55 @@
+#### Preamble ####
+# Purpose: Models the relationship between different economic indicators and vehicle thefts 
+#          using regression analysis to identify significant predictors.
+# Author: Siqi Fei
+# Date: 05 April 2024
+# Contact: fermi.fei@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The script requires the 'tidyverse', 'lubridate', and 'caret' packages for data 
+#                 manipulation, date handling, and machine learning modeling.
+# Additional Notes: This script assumes the data has been preprocessed and is ready for modeling.
+
+
+#### Workspace setup ####
+#install.packages("rstanarm")
+#install.packages("tidyverse")
+#install.packages("rstan")
+#install.packages("modelsummary")
+library(tidyverse)
+library(rstanarm)
+library(rstan)
+library(modelsummary)
+
+#### Read data ####
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+
+### Model data ####
+# Bayesian multiple linear regression model including all CPI categories
+bayesian_linear_model <- stan_glm(Total_Vehicle_Thefts ~ All_items + 
+                                    Recreation_education_and_reading + 
+                                    Goods + Household_operations_furnishings_equipment + 
+                                    Health_and_personal_care +
+                                    Transportation + 
+                                    Shelter + Services + Food + Alcoho_tobacco_cannabis +
+                                    Energy + Gasoline + Clothing_footwear,
+                                  data = analysis_data,
+                                  family = gaussian(),  # Assuming the dependent variable is continuous
+                                  prior = normal(0, 2.5, autoscale = TRUE),  # Normal priors for coefficients
+                                  prior_intercept = normal(0, 2.5, autoscale = TRUE),  # Normal prior for the intercept
+                                  chains = 4, iter = 2000)
+
+# Summarize the model to view coefficients and diagnostics
+print(summary(bayesian_linear_model))
+
+# Diagnostic plots for the Bayesian regression model
+plot(bayesian_linear_model)
+
+# If you need to create a more detailed summary with 'modelsummary'
+modelsummary(bayesian_linear_model)
+
+#### Save model ####
+saveRDS(
+  bayesian_linear_model,
+  file = "models/bayesian_linear_model.rds"
+)
+

### 2024-04-19 paper/paper.qmd
update 3.0 version of paper

@@
 bibliography: references.bib
+frontsize: 10pt
 ---
@@
 
-## Data Measurement
+## Data Cleaning
 
@@
 #| label: tbl-analysisData
-#| tbl-cap: Sample of Analysis Dataset
+#| tbl-cap: Sample of Analysis Dataset (includes only three CPI categories)
 #| echo: false
@@
 selected_data <- cleaned_data[, c(1, 7, 3, 6, ncol(cleaned_data))]
-kable(head(selected_data), caption = "Sample of Analysis Dataset")
+kable(head(selected_data), caption = "Sample of Analysis Dataset (includes only three CPI categories)")
 ```
 
-## Data Distribution
+## Data Measurement
 
@@
 ::: {#fig-CPIDistribution}
-![](images/box_all_cpi.png){width="73%"}
+![](images/box_all_cpi.png){width="75%"}
 
@@
 
-## Model set-up
+## Model Set-up
 
@@
 
-## Model justification
+## Model Justification
 
@@
 ::: {#fig-TimeSeries}
-![](images/Time.png)
+![](images/Time.png){width="110%"}
 
@@
 
-The @tbl-CPIData shows a sample dataset of CPI from @statscanCPI2023. The @tbl-AutoData shows a sample dataset of Vehicle Thetfs Dataset from @torontoPoliceAutoTheft2023.
+The @tbl-CPIData shows a sample dataset of CPI from @statscanCPI2023, it shows all the CPI categories in the dataset, while only have 3 observations.
+The @tbl-AutoData shows a sample dataset of Vehicle Thetfs Dataset from @torontoPoliceAutoTheft2023.
 

### 2024-04-19 paper/paper.qmd
final version of paper

@@
 abstract: "In this study, we researched the monthly incidence of vehicle thefts and inflation from 2014 to 2024 by using data sourced from the Toronto Police Service and Statistics Canada.
-We analyzed the data distribution and characteristics of the two datasets and combined both to conduct a correlation analysis by using Bayesian multiple linear regression model and a Shiny application.
-We found a significant positive correlation between the overall CPI and vehicle theft rates.
-This finding suggests that higher inflation may be associated with an increased number of vehicle theft.
+We analyzed the data distribution and characteristics of the two datasets then combined both to conduct a correlation analysis by using Bayesian multiple linear regression model and a Shiny application.
+A significant positive correlation was found between the overall CPI and vehicle theft rates which suggests that higher inflation may be associated with an increased number of vehicle theft.
 This study highlights the importance of economic indicators in predicting criminal behaviour which in terms provide valuable insights for policymakers and economic analysts."
@@
 These factors have contributed to a sharp rise in inflation rates across Canada.
-Data from @statcanCPI indicates that these changes have made it increasingly difficult for people to afford their daily expenses.
+Data from Statistics Canada 2024 [@statcanCPI] indicates that these changes have made it increasingly difficult for people to afford their daily expenses.
 Meanwhile, regional crime rates have also significantly increased.
-According to the Canada Newswire @equite2023, auto theft trends in Ontario rose dramatically by 48.2%, with a vehicle now stolen every five minutes between 2021 and 2023.
+According to the Canada Newswire [@equite2023], auto theft trends in Ontario rose dramatically by 48.2%, with a vehicle now stolen every five minutes between 2021 and 2023.
 
-@McIntoshLi2012 highlight a critical gap in understanding the economic efficiency of crime prevention programs within the Canadian context, emphasizing the need for evidence-based approaches in crime prevention.
+Previous study [@McIntoshLi2012] highlights a critical gap in understanding the economic efficiency of crime prevention programs within the Canadian context, emphasizing the need for evidencebased approaches in crime prevention.
 
@@
 Employing Bayesian multiple linear regression model, we analyzed the trends and variances in CPI across different sectors and correlated these economic indicators with the frequency of auto thefts in Toronto.
-The datasets were sourced from the Toronto Police Service @torontoPoliceAutoTheft2023 and Statistics Canada @statscanCPI2023.
+The datasets were sourced from the Toronto Police Service [@torontoPoliceAutoTheft2023] and Statistics Canada [@statscanCPI2023].
 Our findings indicate that while an increase in the overall cost of living correlates with higher vehicle theft rates, the rise in Shelter CPI appears to inversely affect theft frequencies.
@@
 
-This paper is structured as follows: @sec-data introduces the datasets utilized for the analysis, offering insights into the variables of interest and their visual representations.
+This paper is structured as follows: @sec-data introduces the datasets utilized for the analysis, offering perspectives into the variables of interest and their visual representations.
 @sec-model details the Bayesian multiple linear regression model applied to explore the relationship between different CPI indicators' change and vehicle theft rates.
@@
 The databases used for this study primarily focus on the Consumer Price Index (CPI) in Canada and vehicle theft incidents in the Toronto area.
-We have utilized open data from @statscanCPI2023 and the @torontoPoliceAutoTheft2023.
-Statistics Canada released the latest update in April 2024 and the data set uses 2002 as the base year to record monthly changes in Canada’s CPI from January 2014 to February 2024 covering 15 different categories.
+We have utilized open data from Statistics Canada [@statscanCPI2023] and the Toronto Police Service [@torontoPoliceAutoTheft2023].
+Statistics Canada released the latest update in April 2024 and the data set uses 2002 as the base year to record monthly changes in Canada’s CPI from January 2014 to February 2024, covering 15 different categories.
 The Toronto Police Service’s Auto Theft Open Data spans from January 2014 to March 2024, detailing each vehicle theft incident with its occurrence time and location.
 
-While other datasets were considered, the CPI data from Statistics Canada was chosen for its authoritative and wide coverage, and the Toronto Police Service's data was selected due to its reliability and detailed record of incidents over a decade.
+While other sources were also available, the CPI data from Statistics Canada was chosen for its authoritative and wide coverage, and the Toronto Police Service's data was selected due to its reliability and detailed record of incidents over a decade.
 These sources provide the most relevant and reliable data for analyzing the relationship between economic conditions and crime rates.
@@
 
-The Consumer Price Index (CPI) data set from Statistics Canada @statscanCPI2023, a sample represented in [Appendix -@sec-dataset-details], utilizes a multi-stage sampling strategy.
+The Consumer Price Index (CPI) dataset from [@statscanCPI2023] utilizes a multi-stage sampling strategy.
+A sample CPI dataset is presented in [Appendix -@sec-dataset-details].
 It targets both urban and rural private households across Canada, while excluding non-representative groups such as inmates or individuals in collective households.
@@
 
-Monthly price data collection relies on detailed specifications for a wide array of goods and services.
-This method allows for accurate tracking of price changes over time, using data from retail outlets and service providers.
+Monthly price data collection relies on detailed specifications for a wide range of goods and services.
+This method allows for accurate tracking of price changes over time.
+The data is sourced from retail outlets and service providers.
 This broad item coverage accurately reflects Canadian consumer spending behaviours.
@@
 
--   `All Items`: an aggregate measure of overall inflation.
--   `Recreation,Education and Reading`: costs related to leisure activities, covering educational expenses and reading materials.
--   `Goods`: including consumer products like electronics and clothing.
--   `Household Operations, Furnishings and Equipment`: expenses for maintaining a household.
--   `Health and Personal Care`: encompassing medical and health-related expenses.
--   `Transportation`: expenses related to personal and public transport.
--   `Shelter`: costs associated with housing.
--   `Services`: various personal and professional services.
--   `Food`: expenses on groceries and dining out.
--   `Alcohol, Tobacco and Cannabis`: spending on alcoholic drinks, tobacco, and cannabis.
--   `Energy`: household energy expenses.
--   `Gasoline`: specifically fuel costs.
--   `Clothing and Footwear`: spending on personal attire.
+-   `All Items`: an aggregate measure of overall inflation
+-   `Recreation,Education and Reading`: costs related to leisure activities, covering educational expenses and reading materials
+-   `Goods`: including consumer products like electronics and clothing
+-   `Household Operations, Furnishings and Equipment`: expenses for maintaining a household
+-   `Health and Personal Care`: encompassing medical and health-related expenses
+-   `Transportation`: expenses related to personal and public transport
+-   `Shelter`: costs associated with housing
+-   `Services`: various personal and professional services
+-   `Food`: expenses on groceries and dining out
+-   `Alcohol, Tobacco and Cannabis`: spending on alcoholic drinks, tobacco, and cannabis
+-   `Energy`: household energy expenses
+-   `Gasoline`: specifically fuel costs
+-   `Clothing and Footwear`: spending on personal attire
 
@@
 
-The Toronto Police Service's dataset on auto theft @torontoPoliceAutoTheft2023 occurrences, a sample represented in [Appendix -@sec-dataset-details], is structured at the offense and vehicle level.
-Each reported occurrence can have multiple records, reflecting different Major Crime Indicators (MCIs) used to categorize each event.
-Data include the report date (REPORT_DATE) and the date the offense occurred (OCC_DATE), both standardized to UTC timezone to ensure uniformity.
-For privacy protection, exact locations of crimes are offset to the nearest road intersection, and the data excludes occurrences deemed unfounded—where an investigation determined that the reported offense did not occur or was not attempted.
+The Toronto Police Service's dataset on auto theft [@torontoPoliceAutoTheft2023] occurrences is structured at the offense and vehicle level, a sample auto thefts dataset is presented in [Appendix -@sec-dataset-details].
+It includes the report date (REPORT_DATE) and the date the offense occurred (OCC_DATE), both standardized to UTC timezone to ensure uniformity.
+For privacy protection, exact locations of crimes are offset to the nearest road intersection.
 
-The auto theft dataset is updated quarterly, reflecting the dynamic nature of crime reporting.
-Additionally, all historical data ranges are provided to ensure coverage of trends over time.
-This dataset is intended to enhance public safety awareness by making detailed crime data accessible to the community while adhering to stringent privacy and data integrity standards.
+The auto theft dataset is updated quarterly, reflecting the dynamic nature of crime reporting. Additionally, all historical data ranges are provided to ensure coverage of trends over time.
 
@@
 
-To ensure a coherent analysis, we synchronized the two datasets by aligning them according to common temporal identifiers, specifically month and year.
+To ensure a coherent analysis, we synchronized the two datasets by aligning them according to month and year.
 The @tbl-analysisData is an example of final analysis dataset after merge CPI dataset and auto thefts dataset.
@@
 
--   Consistency and Alignment Checks: We cross-validated the monthly CPI values against the reported vehicle thefts to ensure that each observation was correctly matched and no temporal misalignments were present.
+-   Consistency and Alignment Checks: We cross-validated the monthly CPI values against the reported vehicle thefts to ensure that each observation was correctly matched and no misalignment were present.
 
-By adhering to these data verification protocols, our study ensures a high level of data quality, enabling a focused and reliable examination of how various economic conditions as depicted by CPI data correlate with trends in vehicle thefts across the Toronto area.
+By using these data verification protocols, our study ensures a high level of data quality, enabling a focused and reliable examination of how various economic conditions as used by CPI data correlate with trends in vehicle thefts across the Toronto area.
 
@@
 #| message: false
+
 # Distribution Plot
@@
 The density curve highlights the overall shape of the distribution, emphasizing the concentration of data points around the mode.
+
+\newpage
 
@@
 ```
-We run the Bayesian model in R [@citeR] using the `rstanarm` package of @rstanarm.
+We run the model in R [@citeR] using the `rstanarm` package of [@rstanarm].
 We apply Normal priors with a mean of zero and standard deviation of 2.5 are applied to each of the coefficients, providing a balance between allowing the data to inform the model and maintaining prior constraints to avoid overly flexible interpretations.
 The error variance $\sigma$ is assumed to follow an exponential distribution, emphasizing the assumption of non-negative error variability.
-This approach captures the nuanced effects of economic indicators on vehicle theft rates, modeling each category's potential impact separately.
+This approach captures the effects of economic indicators on vehicle theft rates, modeling each category's potential impact separately.
 
@@
 
-We expect a notable relationship between the various CPI categories and the rates of vehicle thefts.
+This study employs Bayesian multiple linear regression because it allows for the integration of prior distributions and is better suited to managing the complexity and variability in the data.
+
+We expect a positive relationship between the various CPI categories and the rates of vehicle thefts.
 In particular, economic factors such as $\beta_1$ (All_items), $\beta_6$ (Transportation), and $\beta_{12}$ (Gasoline) may significantly influence theft incidents due to their impact on the cost of living and mobility.
@@
 
-The model results are summarized in @tbl-modelresults.
-
 ```{r}
@@
 
-Our analysis presents a robust relationship between CPI categories and vehicle thefts.
-The model, adjusting for multiple variables, offers a nuanced understanding of how different economic indicators influence theft rates.
+Our analysis presents a positive relationship between CPI categories and vehicle thefts.
+The model, adjusting for multiple variables, offers a better understanding of how different economic indicators influence theft rates.
+The results are summarized in @tbl-modelresults.
 
--   The intercept at -3458.80 establishes a baseline for the model; without any CPI influence, the model predicts a significantly negative number of thefts, which isn't possible in reality and suggests the model's reliance on the CPI predictors.
+-   The intercept at -3458.80 establishes a baseline for the model, without any CPI influence, the model predicts a significantly negative number of thefts, which isn't possible in reality and suggests the model's reliance on the CPI predictors.
 -   The 'All_items' category shows a positive coefficient (52.25), indicating that a unit increase in this category is associated with an increase of approximately 52 vehicle thefts, signifying the strong effect of overall cost of living on theft incidents.
@@
 Our analysis considered multiple CPI categories to identify which aspects of economic change have an effect on crime.
+As evidenced by the results, financial hardship could drive an uptick in thefts, while increased leisure and educational spending seems to reduce them.
 
@@
 
-The pandemic brought significant changes to the Consumer Price Index (CPI), showcasing the economic shifts caused by global lockdowns and altered consumer behaviors.
-Notably, the 'All_items' CPI increased due to heightened demand and supply chain interruptions, escalating the cost of daily goods and services.
-Price fluctuations in the 'Energy' and 'Gasoline' sectors were marked by sharp volatility, reflecting decreased global demand and the impacts of oil price conflicts.
-Conversely, expenditures on 'Recreation_education_and_reading' remained stable, likely due to a shift in spending towards home-based activities during public restrictions.
+In @sec-results, we discussed the time series changes of different CPI categories.
+One notable trend is the unusual shifts in the Consumer Price Index (CPI) and vehicle theft rates during the pandemic period from 2019 to 2021.
+The pandemic significantly impacted the CPI, illustrating how global lockdowns and shifts in consumer behaviour affected the economy.
+The 'All_items' CPI rose due to increased demand and disruptions in supply chains, which escalated the prices of everyday goods and services.
+The 'Energy' and 'Gasoline' sectors experienced big price fluctuations due to decreased global demand and oil price conflicts.
+In contrast, spending on 'Recreation_education_and_reading' remained stable, likely because people invested more in home-based activities during lockdowns.
 
-Simultaneously, the pandemic's economic stress led to an increase in vehicle thefts, mirroring the surge in CPI.
-This rise particularly impacted categories critical for daily living, suggesting that economic pressures may drive individuals towards criminal activities as a coping mechanism.
-Changes in public movement and lockdown measures reshaped traditional crime patterns, indicating shifts in the timing and locations of criminal activities.
-
-This period of profound economic uncertainty and enforced social restrictions clearly demonstrates the interconnectedness of economic factors and crime rates over time.
-It underscores the necessity for vigilant, responsive social and economic strategies during crises to effectively mitigate the escalation of crime, reflecting the enduring impact of such global events on societal norms and safety.
+At the same time, the economic strain from the pandemic led to a significant increase in vehicle thefts, mirroring the rise in the CPI.
+Vehicle thefts spiked notably in Toronto, reflecting trends observed in several urban areas during the pandemic, as discussed in the CBC News article [@cbc2020cartheft].
+This surge was particularly evident in categories essential for daily life, suggesting that financial difficulties may push some individuals toward crime.
+Similar trends have been observed globally, a study by [@northeastern2020crime] details how the pandemic has influenced crime rates across the United States, noting a significant increase in violent crimes such as homicides and assaults during 2020.
+Lockdowns and changes in public behaviour have altered typical crime patterns, affecting when and where crimes occurred.
 
@@
 
-This study also explores the connection between Consumer Price Index (CPI) categories and vehicle theft rates in Canada over a decade.
-It finds that increases in the 'All_items' CPI, reflecting the overall cost of living, align with more vehicle thefts.
-However, spending in areas like recreation and education correlates with fewer thefts, suggesting discretionary spending may deter crime.
+While discussing the pandemic's impact, this study also examines the relationship between various CPI categories and vehicle theft rates in Toronto over the past ten years.
+Similar to findings from research on the economic impacts of the fracking boom [@cgoFracking2020], our analysis indicates complex interactions between economic conditions and crime rates.
+We found that increases in the 'All_items' CPI correspond with higher vehicle theft rates.
+This suggests that as the cost of basic goods and services increases, so does the frequency of vehicle thefts, potentially as individuals face greater financial stress.
+Conversely, we observe that higher spending in non-essential areas like recreation and education is associated with fewer thefts, suggesting that such discretionary expenditures might help reduce crime.
 
-The relationship between vehicle thefts and essential goods like 'Health_and_personal_care' and 'Clothing_footwear' appears less direct.
-Meanwhile, 'Energy' and 'Gasoline' prices show unique patterns, indicating market fluctuations might influence theft rates differently.
+Additionally, while the initial hypothesis of this study suggested a strong correlation between 'Energy' and 'Gasoline' CPI and the frequency of vehicle thefts, the results displayed only a weak association (coefficients of 0.99 and 2.17 respectively).
+According to a CTV News report [@ctvnews2022gasanddash], surges in 'Energy' and 'Gasoline' CPI tend to coincide more with increases in gas theft rather than vehicle thefts.
+This indicates that economic strains may prompt shifts in the type of crimes committed, with individuals potentially opting for more immediate criminal opportunities such as gas theft.
 
-By examining CPI categories separately, the study highlights how economic conditions affect crime.
-It offers insights that could guide policies for economic management and crime prevention, showing how various economic pressures impact crime differently and underscoring the role of economic analysis in improving societal conditions.
+The findings provide valuable information that could help shape economic and crime prevention strategies, demonstrating how different types of economic stress affect crime rates and highlighting the importance of economic analysis in societal improvement.
 
@@
 
-This study aimed to explore the impact of economic changes on crime rates. However, it faces several limitations:
+The primary limitations of this study are mainly focused on those perspectives:
 
 1.
-**Scope of Economic Factors:** The research focused on CPI and vehicle thefts but did not include other potentially influential economic factors such as unemployment rates or income disparity.
-Expanding the range of economic indicators could provide a broader understanding of what drives crime rates.
+**Limited Economic Factors**: The analysis was restricted to CPI and vehicle thefts, excluding other significant economic factors such as unemployment rates or income disparity.
+Including more economic indicators could enhance our understanding of the drivers behind crime rates.
 
 2.
-**Geographic Limitation:** The findings are based on data from Toronto and may not apply to other regions with different socio-economic conditions.
-Future studies could analyze data from multiple regions to draw more general conclusions.
+**Geographic Limitation**: The results are based on Toronto data and may not be representative of other areas with different socio-economic backgrounds.
+Future research should include data from various geographic locations to validate the findings more broadly.
 
 3.
-**Population Bias:** As a major urban center, Toronto's large population might influence the study results, potentially making them less applicable to smaller or rural areas.
-Including a more varied set of locations could help balance this bias.
+**Population Bias**: The study focuses on Toronto, a large urban center, which may not reflect the conditions in smaller or rural areas.
+A broader geographic sample could provide a more balanced view.
 
 4.
-**Impact of the Pandemic:** The study acknowledges the pandemic's effects but does not analyze them in depth.
-Further research could more closely examine how specific aspects of the pandemic influenced economic and crime trends.
+**Pandemic Analysis**: While the study notes the impact of the pandemic, it does not delve deeply into its specific effects on economic and crime trends.
+More detailed research on this topic could yield valuable insights.
 
 5.
-**Data Reliability:** The study relies on reported crime data and CPI, which could be subject to underreporting or inconsistencies in data collection.
-More rigorous data verification could strengthen the findings.
+**Data Reliability**: The study relies on reported crime data and CPI, which might not be entirely accurate due to potential underreporting or data collection inconsistencies.
+Enhancing data verification processes could improve the reliability of the findings.
 
 6.
-**Causal Pathways:** The study does not investigate how economic changes directly lead to variations in crime rates.
-Future research might include qualitative approaches to better understand the causal relationships involved.
+**Causal Relationships**: The research does not explore the direct causal mechanisms linking economic changes to crime rates.
+Future studies could benefit from incorporating qualitative research to better understand these relationships.
 
@@
 
-Given the limitations identified in the study, here are what we can do in the future study:
+Given the limitations identified in the study, here are what we can improve in the future studies:
 
 1.
-**Broader Economic Indicators**: Future studies should incorporate a wider range of economic indicators, including unemployment rates, income levels, and economic policy changes, to more fully understand their impact on crime rates.
+**Broader Economic Indicators**: Future studies should expand the range of economic indicators examined to include factors like unemployment rates and income levels, which could provide a more comprehensive view of their impact on crime rates.
 
 2.
-**Expanded Geographic Scope**: To address the geographic limitation, subsequent research could include data from multiple cities and rural areas across different countries.
-This would help determine if the findings from Toronto are applicable elsewhere or if local conditions significantly influence crime dynamics.
+**Expanded Geographic Scope**: Future research should analyze data from multiple cities and rural areas across various countries to check if the findings in Toronto apply to other areas or if local conditions significantly alter crime dynamics.
 
 3.
-**Diverse Demographic Analysis**: Investigating how economic changes affect different demographic groups could uncover nuanced patterns of how economic conditions influence crime rates among various populations.
+**Diverse Demographic Analysis**: Examining how economic changes affect different demographic groups could reveal more detailed patterns of how economic conditions influence crime rates across various communities.
 
-4.
-**In-depth Pandemic Analysis**: A focused study on the pandemic period could clarify how temporary economic disruptions and social restrictions specifically affected crime rates, providing insights into managing crime during crises.
+4.  **In-depth Pandemic Analysis**: A focused analysis of the economic disruptions and social restrictions during the pandemic could better explain how these factors specifically impacted crime rates.
 
 5.
-**Enhanced Data Collection and Verification**: To improve data reliability, future research should use more rigorous methods for data collection and verification.
-This could involve cross-referencing crime data with other sources or using advanced statistical techniques to handle potential data inaccuracies.
+**Enhanced Data Collection and Verification**: Future studies should employ more stringent data collection and verification methods, possibly by cross-referencing crime data with other sources or utilizing advanced statistical techniques to address potential inaccuracies.
 
 6.
-**Causal Research**: Implementing mixed-methods approaches, including qualitative research, could help elucidate the causal pathways between economic conditions and crime, offering a clearer picture of the underlying mechanisms.
+**Causal Research**: Incorporating mixed-methods approaches and qualitative research could clarify the causal pathways between economic conditions and crime, offering a clearer understanding of the underlying mechanisms.
 
 7.
-**Policy Impact Studies**: Researching the effects of specific economic and social policies on crime rates could provide valuable feedback for policymakers, helping to design interventions that effectively address the economic drivers of crime.
-
-By pursuing these directions, future research can build on the current study’s findings and address its shortcomings, leading to a deeper understanding of the complex relationship between economics and crime.
+**Policy Impact Studies**: Analyzing the effects of specific economic and social policies on crime rates could help policymakers design more effective interventions to address the economic drivers of crime.
 
@@
 
-The @tbl-CPIData shows a sample dataset of CPI from @statscanCPI2023, it shows all the CPI categories in the dataset, while only have 3 observations.
+The @tbl-CPIData shows a sample dataset of CPI from @statscanCPI2023, it has all the CPI categories in the dataset, while only have 3 observations.
 The @tbl-AutoData shows a sample dataset of Vehicle Thetfs Dataset from @torontoPoliceAutoTheft2023.
@@
 
-The @fig-LPAll-items shows a sample scenario in the shiny app. The shiny app can be found through link: <https://siqi-fei.shinyapps.io/VehicleTheft_Inflation/>
+The @fig-LPAll-items and @fig-shiny show a sample scenario in the shiny app. The shiny app can be found through link: <https://siqi-fei.shinyapps.io/VehicleTheft_Inflation/>
 
@@
 ```
+
+::: {#fig-shiny}
+![](images/Shiny App.png){width="80%"}
+
+Sample of Shiny App
+:::
 
@@
 
+\newpage
+
 ## Posterior predictive check

### 2024-04-19 Vehicle_Theft_Shiny_App/app.R

@@ -1,26 +1,44 @@
+#### Preamble ####
+# Purpose: This script creates an interactive data viewer for analyzing the 
+#          relationship between CPI categories and vehicle theft rates in 
+#          Toronto. It allows users to select different CPI categories and
+#          visualize their correlation with vehicle theft occurrences over time.
+# Author: Siqi Fei
+# Date: 05 April 2024
+# Contact: fermi.fei@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: R environment with the following packages installed: `shiny`, 
+#                 `ggplot2`, `DT`.
+# Additional Notes: This application is intended to provide a user-friendly 
+#                   interface for dynamic exploration of the dataset, enabling
+#                   both visualization and data download capabilities. Ensure
+#                   the dataset "analysis_data_copy.csv" is updated and located
+#                   in the same directory as this script for optimal performance.
+
+
 #### Work space Setup ####
-library(shiny)
-library(ggplot2)
-library(DT)
+library(shiny)          # Load shiny for interactive web apps
+library(ggplot2)        # Load ggplot2 for creating visualizations
+library(DT)             # Load DT for rendering interactive tables
 
 # Load dataset
 transposed_data <- read.csv("analysis_data_copy.csv")
 
 # Define the user interface
 ui <- fluidPage(
-  titlePanel("Interactive Data Viewer"),
+  titlePanel("Interactive Toronto CPI and Auto Thefts Viewer"), # Title of the app
   sidebarLayout(
     sidebarPanel(
-      selectInput("xAxisColumn", "Select CPI Category for X-axis:",
+      selectInput("xAxisColumn", "Select CPI Category for X-axis:", # Dropdown menu for selecting CPI category
                   choices = colnames(transposed_data)[-which(colnames(transposed_data) == "Total_Vehicle_Thefts")],
                   selected = "All_items"),
-      hr(),
-      downloadButton("downloadData", "Download Data")
+      hr(), # Horizontal line for layout separation
+      downloadButton("downloadData", "Download Data") # Button to download the data
     ),
     mainPanel(
-      DTOutput("dataTable"),
-      hr(),
-      plotOutput("dataPlot")
+      DTOutput("dataTable"), # Output table display
+      hr(), # Horizontal line for layout separation
+      plotOutput("dataPlot") # Output plot display
     )
   )
 )
[…]

### 2024-04-19 scripts/00-simulate_data.R

@@ -96,8 +96,6 @@ for (col in names(simulate_data)[4:ncol(simulate_data)]) {
     scale_y_continuous(sec.axis = sec_axis(~., name = "CPI"), name = "Auto Theft Incidents")
 }
 
-# Show the plot
-print(auto_theft_plot)
 
 #### Save plot ####
 # Saving the plot

### 2024-04-19 scripts/01-data_cleaning.R

@@ -85,74 +85,74 @@ names(vehicle_theft_wide) <- sapply(names(vehicle_theft_wide), convert_to_month_
 
 # Merge Dataset
 merge_data <- merge(raw_data_cpi, vehicle_theft_wide, by.x = c("January 2014",
-                    "February 2014","March 2014","April 2014","May 2014",
-                    "June 2014","July 2014","August 2014","September 2014",
-                    "October 2014","November 2014","December 2014",
-                    "January 2015","February 2015","March 2015","April 2015",
-                    "May 2015","June 2015","July 2015","August 2015",
-                    "September 2015","October 2015","November 2015",
-                    "December 2015","January 2016","February 2016","March 2016",
-                    "April 2016","May 2016","June 2016","July 2016",
-                    "August 2016","September 2016","October 2016",
-                    "November 2016","December 2016","January 2017",
-                    "February 2017","March 2017","April 2017","May 2017",
-                    "June 2017","July 2017","August 2017","September 2017",
-                    "October 2017","November 2017","December 2017",
-                    "January 2018","February 2018","March 2018","April 2018",
-                    "May 2018","June 2018","July 2018","August 2018",
-                    "September 2018","October 2018","November 2018",
-                    "December 2018","January 2019","February 2019","March 2019",
-                    "April 2019","May 2019","June 2019","July 2019",
-                    "August 2019","September 2019","October 2019",
-                    "November 2019","December 2019","January 2020",
-                    "February 2020","March 2020","April 2020","May 2020",
-                    "June 2020","July 2020","August 2020","September 2020",
-                    "October 2020","November 2020","December 2020",
-                    "January 2021","February 2021","March 2021","April 2021",
-                    "May 2021","June 2021","July 2021","August 2021",
-                    "September 2021","October 2021","November 2021",
-                    "December 2021","January 2022","February 2022","March 2022",
-                    "April 2022","May 2022","June 2022","July 2022",
-                    "August 2022","September 2022","October 2022",
-                    "November 2022","December 2022","January 2023",
-                    "February 2023","March 2023","April 2023","May 2023",
-                    "June 2023","July 2023","August 2023","September 2023",
-                    "October 2023","November 2023","December 2023",
-                    "January 2024","February 2024"), by.y = c("January 2014",
-                    "February 2014","March 2014","April 2014","May 2014",
-                    "June 2014","July 2014","August 2014","September 2014",
-                    "October 2014","November 2014","December 2014",
[…]

### 2024-04-19 scripts/02-test_data.R

@@ -27,7 +27,6 @@ transposed_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
 raw_data_cpi <- read.csv(here::here("data/analysis_data/analysis_data_CPI.csv"))
 vehicle_theft_wide <- read.csv(here::here("data/analysis_data/analysis_data_Vehicle.csv"))
 
-
 #### Test data ####
 
 #### Tests for CPI Data Cleaning ####
@@ -73,4 +72,6 @@ test_that("Merged data has no NA values", {
 test_that("CPI data contains only numeric columns", {
   expect_true(all(sapply(transposed_data[-1], is.numeric)))
 })
+
+
 

### 2024-04-19 scripts/03-data_measurement.R

@@ -9,7 +9,9 @@
 #                 date handling. Ensure that the raw data is in the specified directory.
 # Additional Notes: This script generates a cleaned dataset saved in a standard CSV format for use in modeling.
 
-#### Set Up ###
+#### Workspace Setup ###
+
+# Load necessary libraries for data manipulation and visualization
 #install.packages("tidyr")
 #install.packages("corrplot")
 #install.packages("lubridate")
@@ -20,13 +22,15 @@ library(corrplot)
 library(lubridate)
 
 #### Load Data ####
+
+# Load datasets from specified paths
 merge_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
 raw_data_cpi <- read.csv(here::here("data/analysis_data/analysis_data_CPI.csv"))
 vehicle_theft_wide <- read.csv(here::here("data/analysis_data/analysis_data_Vehicle.csv"))
 
 #### CPI Data Measurements ####
 
-# Pivot the data to long format for plotting
+# Transform CPI data from wide to long format to facilitate plotting
 data_long <- pivot_longer(merge_data, 
                           cols = c(Recreation_education_and_reading, Goods, Household_operations_furnishings_equipment,
                                    Health_and_personal_care, All_items, Transportation, Shelter, Services, Food,
@@ -34,7 +38,7 @@ data_long <- pivot_longer(merge_data,
                           names_to = "Category", 
                           values_to = "Index")
 
-# Scatter Plot with Regression Line
+# Create scatter plots with regression lines for CPI categories vs vehicle thefts
 ggplot(merge_data, aes(x = All_items, y = Total_Vehicle_Thefts)) +
   geom_point() +
   geom_smooth(method = "lm", se = FALSE) +
@@ -49,7 +53,7 @@ ggplot(merge_data, aes(x = Energy, y = Total_Vehicle_Thefts)) +
        x = "Energy CPI",
        y = "Total Vehicle Thefts")
 
-# Exclude All_items and Energy CPI categories
+# Excluding specific CPI categories for focused analysis
 merge_data_filtered <- merge_data[, !names(merge_data) %in% c("All_items", "Energy")]
 
 # Plot linear regression for each CPI category
@@ -98,7 +102,7 @@ legend("topright",
        bty = "n")
 
 # Add a rug plot to show individual data points
-rug(merge_data$Total_Vehicle_Thefts, col = "darkblue", lwd = 0.5)
[…]

### 2024-04-19 scripts/04-model.R

@@ -5,7 +5,7 @@
 # Date: 05 April 2024
 # Contact: fermi.fei@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The script requires the 'tidyverse', 'lubridate', and 'caret' packages for data 
+# Pre-requisites: The script requires the `rstanarm`, 'tidyverse', 'lubridate', and 'caret' packages for data 
 #                 manipulation, date handling, and machine learning modeling.
 # Additional Notes: This script assumes the data has been preprocessed and is ready for modeling.
 
@@ -44,7 +44,7 @@ print(summary(bayesian_linear_model))
 # Diagnostic plots for the Bayesian regression model
 plot(bayesian_linear_model)
 
-# If you need to create a more detailed summary with 'modelsummary'
+# create a more detailed summary with 'modelsummary'
 modelsummary(bayesian_linear_model)
 
 #### Save model ####

### 2024-04-19 paper/paper.qmd
fixed typo in final version

@@
 These factors have contributed to a sharp rise in inflation rates across Canada.
-Data from Statistics Canada 2024 [@statcanCPI] indicates that these changes have made it increasingly difficult for people to afford their daily expenses.
+Data from Statistics Canada [@statcanCPI] indicates that these changes have made it increasingly difficult for people to afford their daily expenses.
 Meanwhile, regional crime rates have also significantly increased.
@@
 This surge was particularly evident in categories essential for daily life, suggesting that financial difficulties may push some individuals toward crime.
-Similar trends have been observed globally, a study by [@northeastern2020crime] details how the pandemic has influenced crime rates across the United States, noting a significant increase in violent crimes such as homicides and assaults during 2020.
+Similar trends have been observed globally, a study by Northeastern University [@northeastern2020crime] details how the pandemic has influenced crime rates across the United States, noting a significant increase in violent crimes such as homicides and assaults during 2020.
 Lockdowns and changes in public behaviour have altered typical crime patterns, affecting when and where crimes occurred.
@@
 While discussing the pandemic's impact, this study also examines the relationship between various CPI categories and vehicle theft rates in Toronto over the past ten years.
-Similar to findings from research on the economic impacts of the fracking boom [@cgoFracking2020], our analysis indicates complex interactions between economic conditions and crime rates.
+Similar to findings from research on the economic impacts of the fracking boom [@cgo2020fracking], our analysis indicates complex interactions between economic conditions and crime rates.
 We found that increases in the 'All_items' CPI correspond with higher vehicle theft rates.
-This suggests that as the cost of basic goods and services increases, so does the frequency of vehicle thefts, potentially as individuals face greater financial stress.
+This shows that as the cost of basic goods and services increases, so does the frequency of vehicle thefts, potentially as individuals face greater financial stress.
 Conversely, we observe that higher spending in non-essential areas like recreation and education is associated with fewer thefts, suggesting that such discretionary expenditures might help reduce crime.
